---
title: "Similarity analysis comparing biological vs random mother-infant pairs for OIF cohorts"
author: "Tiange Liu"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/USERS/TiangeLiu/OUTPUT/Results",  
    )
  })
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
source("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/USERS/TiangeLiu/ANALYSIS/Codes/OIF_Function.r")
path <- paste("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/USERS/TiangeLiu/OUTPUT/")
```


Read in covariates 
```{r}
master_dt <- read.csv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/OIF_master_SQL_072823_v3.csv")

# re-label some variables
master_dt$birth_delivm <- as.factor(master_dt$birth_delivm)
master_dt$birth_delivm <- factor(master_dt$birth_delivm, levels=c("-5", "1","2"), labels=c("Missing", "Vaginal delivery", "Cesarean delivery"))

master_dt$med_antib_pn <- as.factor(master_dt$med_antib_pn)
master_dt$med_antib_pn <- factor(master_dt$med_antib_pn, levels=c("-8", "0", "1"), labels=c("Missing", "No antibiotics", "Took antibiotics"))
```


# MARCH
## Prepare master data
```{r message=FALSE}
## Read in paired IDs
march <- read.csv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/12902/MARCH_paired_54.csv")

march$participant_type <- as.factor(march$m16s_part_type)
march$participant_type <- factor(march$participant_type, levels=c("1", "2"), labels=c("biological mother", "child"))

march$body_site <- as.factor(march$m16s_spec_type)
march$body_site <- factor(march$m16s_spec_type, levels=c("1", "7"), labels=c("stool", "vaginal"))

march$paired_ID <- substr(march$ParticipantID, 1, 9)


## Add breastfeeding variables measured at the time of sample collection
bf_dt <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/01.metadata/submitted_participants_0556a.tsv")

table(bf_dt$child_breastfed, bf_dt$infant_diet)
# breastfed: 1=yes, 2=no
# diet:1, Breastmilk| 2, Infant formula | 3, Mixed breastmilk and formula |4, Breastmilk and solids| 5, Formula and solids| 6, Breastmilk, formula and solids|7, Other| -7,Declined | -8,Don’t know | -9,Missing 

bf_dt <- bf_dt %>%
  select('sample-id', child_breastfed, infant_diet)
march <- merge(march, bf_dt, by.x=c("sample.id"), by.y="sample-id")
table(march[march$participant_type=="child",]$child_breastfed, march[march$participant_type=="child",]$infant_diet)

# generate a new variable
march$feeding <- ifelse(march$participant_type=="biological mother", -6, 
                        ifelse(march$infant_diet==1, 1, ifelse(march$infant_diet==2 & march$child_breastfed==2, 2, 3)))
table(march[march$participant_type=="child",]$feeding)  #34, 6, 14
march$feeding <- as.factor(march$feeding)
march$feeding <- factor(march$feeding, levels=c("-6", "1", "2", "3"), labels=c("NA", "Exclusive breastmilk", "Exclusive formula", "Mixed"))


march$bf <- ifelse(march$participant_type=="biological mother", -6,
                   ifelse(march$feeding=="Exclusive formula", 2, 1))
table(march[march$participant_type=="child",]$bf)  #48, 6
march$bf <- as.factor(march$bf)
march$bf <- factor(march$bf, levels=c("-6", "1", "2"), labels=c("NA", "Yes", "No"))

## Merge with covariates
march <- merge(march, master_dt[, -c(1)], by.x="paired_ID", by.y="PregID", all=FALSE)

## Exclude infant with missing delivery mode
march <- filter(march, !is.na(birth_delivm))

## Restrict to 3rd trimester maternal samples
march <- filter(march, paired_ID %in% as.list(march[march$body_site=="vaginal" & march$m16s_coll_tmst==3,]$paired_ID))

length(unique(march$sample.id))
length(unique(march$paired_ID))
table(march$m16s_age_yr)
```

## Extract infant ID, maternal vaginal ID, maternal stool ID
```{r}
i_id <- march %>% filter(participant_type=="child") %>% select(sample.id)
mv_id <- march %>% filter(body_site=="vaginal") %>% select(sample.id)
ms_id <- march %>% filter(body_site=="stool" & participant_type=="biological mother") %>% select(sample.id)
```

## Read in dissimilarity matrices
```{r message=FALSE, warning=FALSE}
# bray-curtis
bc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12902/braycurtis.tsv") 
row_name <- bc$...1
bc <- bc[, c(-1)]
row.names(bc) <- row_name

# jaccard
jc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12902/jaccard.tsv")
row_name <- jc$...1
jc <- jc[, c(-1)]
row.names(jc) <- row_name

# unweighted unifrac
uu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12902/unweighted_unifrac.tsv")
row_name <- uu$...1
uu <- uu[, c(-1)]
row.names(uu) <- row_name

# weighted unifrac
wu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12902/weighted_unifrac.tsv")
row_name <- wu$...1
wu <- wu[, c(-1)]
row.names(wu) <- row_name
```

## Create summary datasets for dissimilarity within pairs of mother-own child, mother-unrelated child
```{r message=FALSE}
## based on rarefied count table
sum_stool_march <- combine_sum_diss(bc, jc, uu, wu, i_id, ms_id)
sum_stool_march$age <- "3 (3, 6)"
sum_stool_march$cohort <- "MARCH"


sum_vaginal_march <- combine_sum_diss(bc, jc, uu, wu, i_id, mv_id)
sum_vaginal_march$age <- "3 (3, 6)"
sum_vaginal_march$cohort <- "MARCH"


## based on un-rarefied count table
asv_table <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12902/12902_asv_table_classic.tsv")
  # this is newer asv table Justine created in May 2023

asv_table <- as.data.frame(t(asv_table)) 
  # rows are samples, columns are taxa
colnames(asv_table) <- asv_table[1, ]
  # assign colnames as ASV id
asv_table <- asv_table[-(1:2), ] # remove unnecessary rows
asv_table <- asv_table[-(nrow(asv_table)), ] # remove unnecessary rows

row_names <- row.names(asv_table)
asv_table <- as.data.frame(lapply(asv_table, as.integer))
asv_table <- as.matrix(asv_table)
row.names(asv_table) <- row_names

jc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "jaccard")))
bc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "bray")))


sum_stool_march_2 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id, ms_id)
sum_stool_march_2$age <- "3 (3, 6)"
sum_stool_march_2$cohort <- "MARCH"


sum_vaginal_march_2 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id, mv_id)
sum_vaginal_march_2$age <- "3 (3, 6)"
sum_vaginal_march_2$cohort <- "MARCH"
```



# Import participant data that will be useful for the other four cohorts
```{r}
full_dta <- read.delim("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/01.metadata/m16s_der_hpc_0556a.tsv")

table(full_dta$m16s_part_type)
full_dta$m16s_part_type <- as.factor(full_dta$m16s_part_type)
full_dta$m16s_part_type <- factor(full_dta$m16s_part_type, levels=c("1", "2"), labels=c("biological mother", "child"))

table(full_dta$m16s_spec_type)
full_dta$body_site <- as.factor(full_dta$m16s_spec_type)
full_dta$body_site <- factor(full_dta$m16s_spec_type, levels=c("1", "7"), labels=c("stool", "vaginal"))
```

# VDAART
## Prepare master data
```{r}
vdaart <- read.csv(paste0(path,"Results/feast_12301_091123.csv"))

table(vdaart$m16s_age_yr)
# possible cut off point: <=0.5y, 0.83-1.25y, 2.83-3.25, 3.92-4.25
```

## Extract infant ID, maternal vaginal ID, maternal stool ID
```{r}
i_id_1 <- vdaart %>% filter(m16s_age_yr<=0.5) %>% select(sample.id)
ms_id_1 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==12301 & 
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_1$sample.id, 7, 15)))) %>% select(sample.id)

i_id_2 <- vdaart %>% filter(m16s_age_yr>=0.83 & m16s_age_yr<=1.25) %>% select(sample.id)
ms_id_2 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==12301 & 
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_2$sample.id, 7, 15)))) %>% select(sample.id)

i_id_3 <- vdaart %>% filter(m16s_age_yr>=2.83 & m16s_age_yr<=3.25) %>% select(sample.id)
ms_id_3 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==12301 & 
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_3$sample.id, 7, 15)))) %>% select(sample.id)

i_id_4 <- vdaart %>% filter(m16s_age_yr>=3.92 & m16s_age_yr<=4.25) %>% select(sample.id)
ms_id_4 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==12301 & 
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_4$sample.id, 7, 15)))) %>% select(sample.id)

```


## Read in dissimilarity matrices
```{r message=FALSE, warning=FALSE}
# bray-curtis
bc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12301/braycurtis.tsv") 
row_name <- bc$...1
bc <- bc[, c(-1)]
row.names(bc) <- row_name

# jaccard
jc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12301/jaccard.tsv")
row_name <- jc$...1
jc <- jc[, c(-1)]
row.names(jc) <- row_name

# unweighted unifrac
uu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12301/unweighted_unifrac.tsv")
row_name <- uu$...1
uu <- uu[, c(-1)]
row.names(uu) <- row_name

# weighted unifrac
wu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12301/weighted_unifrac.tsv")
row_name <- wu$...1
wu <- wu[, c(-1)]
row.names(wu) <- row_name
```


## Create summary datasets for dissimilarity within pairs of mother-own child, mother-unrelated child
```{r message=FALSE}
## based on rarefied count table
sum_stool_1 <- combine_sum_diss(bc, jc, uu, wu, i_id_1, ms_id_1)
sum_stool_1$age <- "4 (2, 6)"

sum_stool_2 <- combine_sum_diss(bc, jc, uu, wu, i_id_2, ms_id_2)
sum_stool_2$age <- "12 (10, 15)"

sum_stool_3 <- combine_sum_diss(bc, jc, uu, wu, i_id_3, ms_id_3)
sum_stool_3$age <- "36 (34, 39)"

sum_stool_4 <- combine_sum_diss(bc, jc, uu, wu, i_id_4, ms_id_4)
sum_stool_4$age <- "48 (47, 51)"

sum_stool_vdaart <- rbind(sum_stool_1, sum_stool_2, sum_stool_3, sum_stool_4)
sum_stool_vdaart$cohort <- "VDAART"


## based on un-rarefied count table
asv_table <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12301/12301_asv_table_classic.tsv")
  # this is newer asv table Justine created in May 2023

asv_table <- as.data.frame(t(asv_table)) 
  # rows are samples, columns are taxa
colnames(asv_table) <- asv_table[1, ]
  # assign colnames as ASV id
asv_table <- asv_table[-(1:2), ] # remove unnecessary rows
asv_table <- asv_table[-(nrow(asv_table)), ] # remove unnecessary rows

row_names <- row.names(asv_table)
asv_table <- as.data.frame(lapply(asv_table, as.integer))
asv_table <- as.matrix(asv_table)
row.names(asv_table) <- row_names

jc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "jaccard")))
bc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "bray")))


sum_stool_1 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_1, ms_id_1)
sum_stool_1$age <- "4 (2, 6)"

sum_stool_2 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_2, ms_id_2)
sum_stool_2$age <- "12 (10, 15)"

sum_stool_3 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_3, ms_id_3)
sum_stool_3$age <- "36 (34, 39)"

sum_stool_4 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_4, ms_id_4)
sum_stool_4$age <- "48 (47, 51)"

sum_stool_vdaart_2 <- rbind(sum_stool_1, sum_stool_2, sum_stool_3, sum_stool_4)
sum_stool_vdaart_2$cohort <- "VDAART"

```


# RESONANCE
```{r}
resonance <- read.csv(paste0(path,"Results/feast_10703_091123.csv"))

table(resonance$m16s_age_yr)
# possible cut-off: 0.08-0.25, 0.33-0.67, 0.75-1.58, 1.67-3.42

```


## Extract infant ID, maternal vaginal ID, maternal stool ID
```{r message=FALSE}
i_id_1 <- resonance %>% filter(m16s_age_yr<=0.25) %>% select(sample.id)
ms_id_1 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==10703 &
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_1$sample.id, 7, 15)))) %>% 
  mutate(prefix = substr(sample.id, 7, 15)) %>%
  group_by(prefix) %>%
  slice_tail(n = 1) %>%   # only keep the last observation for each mother
  select(sample.id)
ms_id_1 <- ms_id_1[, c(-1)]


i_id_2 <- resonance %>% filter(m16s_age_yr>=0.33 & m16s_age_yr<=0.67) %>% select(sample.id)
ms_id_2 <- full_dta %>% filter(m16s_part_type=="biological mother" &  CohortID==10703 &
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_2$sample.id, 7, 15)))) %>% 
  mutate(prefix = substr(sample.id, 7, 15)) %>%
  group_by(prefix) %>%
  slice_tail(n = 1) %>%   # only keep the last observation for each mother
  select(sample.id)
ms_id_2 <- ms_id_2[, c(-1)]


i_id_3 <- resonance %>% filter(m16s_age_yr>=0.75 & m16s_age_yr<=1.58) %>% select(sample.id)
ms_id_3 <- full_dta %>% filter(m16s_part_type=="biological mother" &  CohortID==10703 &
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_3$sample.id, 7, 15)))) %>% 
  mutate(prefix = substr(sample.id, 7, 15)) %>%
  group_by(prefix) %>%
  slice_tail(n = 1) %>%   # only keep the last observation for each mother
  select(sample.id)
ms_id_3 <- ms_id_3[, c(-1)]


i_id_4 <- resonance %>% filter(m16s_age_yr>=1.67 & m16s_age_yr<=3.42) %>% select(sample.id)
ms_id_4 <- full_dta %>% filter(m16s_part_type=="biological mother" &  CohortID==10703 &
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_4$sample.id, 7, 15)))) %>% 
  mutate(prefix = substr(sample.id, 7, 15)) %>%
  group_by(prefix) %>%
  slice_tail(n = 1) %>%   # only keep the last observation for each mother
  select(sample.id)
ms_id_4 <- ms_id_4[, c(-1)]
```

## Read in dissimilarity matrices
```{r message=FALSE, warning=FALSE}
# bray-curtis
bc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/10703/braycurtis.tsv") 
row_name <- bc$...1
bc <- bc[, c(-1)]
row.names(bc) <- row_name

# jaccard
jc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/10703/jaccard.tsv")
row_name <- jc$...1
jc <- jc[, c(-1)]
row.names(jc) <- row_name

# unweighted unifrac
uu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/10703/unweighted_unifrac.tsv")
row_name <- uu$...1
uu <- uu[, c(-1)]
row.names(uu) <- row_name

# weighted unifrac
wu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/10703/weighted_unifrac.tsv")
row_name <- wu$...1
wu <- wu[, c(-1)]
row.names(wu) <- row_name
```


## Create summary datasets for dissimilarity within pairs of mother-own child, mother-unrelated child
```{r message=FALSE}
## based on rarefied count table
sum_stool_1 <- combine_sum_diss(bc, jc, uu, wu, i_id_1, ms_id_1)
sum_stool_1$age <- paste0(round(mean(resonance[resonance$m16s_age_yr<=0.25,]$m16s_age_yr)*12,0), " (",
                          round(min(resonance[resonance$m16s_age_yr<=0.25,]$m16s_age_yr)*12,0), ", ",
                          round(max(resonance[resonance$m16s_age_yr<=0.25,]$m16s_age_yr)*12,0), ")")

sum_stool_2 <- combine_sum_diss(bc, jc, uu, wu, i_id_2, ms_id_2)
sum_stool_2$age <- paste0(round(mean(resonance[resonance$m16s_age_yr>=0.33 & resonance$m16s_age_yr<=0.67,]$m16s_age_yr)*12,0), " (",
                          round(min(resonance[resonance$m16s_age_yr>=0.33 & resonance$m16s_age_yr<=0.67,]$m16s_age_yr)*12,0), ", ",
                          round(max(resonance[resonance$m16s_age_yr>=0.33 & resonance$m16s_age_yr<=0.67,]$m16s_age_yr)*12,0), ")")

sum_stool_3 <- combine_sum_diss(bc, jc, uu, wu, i_id_3, ms_id_3)
sum_stool_3$age <- paste0(round(mean(resonance[resonance$m16s_age_yr>=0.75 & resonance$m16s_age_yr<=1.58,]$m16s_age_yr)*12,0), " (",
                          round(min(resonance[resonance$m16s_age_yr>0.75 & resonance$m16s_age_yr<=1.58,]$m16s_age_yr)*12,0), ", ",
                          round(max(resonance[resonance$m16s_age_yr>=0.75 & resonance$m16s_age_yr<=1.58,]$m16s_age_yr)*12,0), ")")

sum_stool_4 <- combine_sum_diss(bc, jc, uu, wu, i_id_4, ms_id_4)
sum_stool_4$age <- paste0(round(mean(resonance[resonance$m16s_age_yr>=1.67 & resonance$m16s_age_yr<=3.42,]$m16s_age_yr)*12,0), " (",
                          round(min(resonance[resonance$m16s_age_yr>=1.67 & resonance$m16s_age_yr<=3.42,]$m16s_age_yr)*12,0), ", ",
                          round(max(resonance[resonance$m16s_age_yr>=1.67 & resonance$m16s_age_yr<=3.42,]$m16s_age_yr)*12,0), ")")

sum_stool_resonance <- rbind(sum_stool_1, sum_stool_2, sum_stool_3, sum_stool_4)
sum_stool_resonance$cohort <- "RESONANCE"


## based on un-rarefied count table
asv_table <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/10703/10703_asv_table_classic.tsv")
  # this is newer asv table Justine created in May 2023

asv_table <- as.data.frame(t(asv_table)) 
  # rows are samples, columns are taxa
colnames(asv_table) <- asv_table[1, ]
  # assign colnames as ASV id
asv_table <- asv_table[-(1:2), ] # remove unnecessary rows
asv_table <- asv_table[-(nrow(asv_table)), ] # remove unnecessary rows

row_names <- row.names(asv_table)
asv_table <- as.data.frame(lapply(asv_table, as.integer))
asv_table <- as.matrix(asv_table)
row.names(asv_table) <- row_names

jc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "jaccard")))
bc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "bray")))

sum_stool_1 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_1, ms_id_1)
sum_stool_1$age <- paste0(round(mean(resonance[resonance$m16s_age_yr<=0.25,]$m16s_age_yr)*12,0), " (",
                          round(min(resonance[resonance$m16s_age_yr<=0.25,]$m16s_age_yr)*12,0), ", ",
                          round(max(resonance[resonance$m16s_age_yr<=0.25,]$m16s_age_yr)*12,0), ")")

sum_stool_2 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_2, ms_id_2)
sum_stool_2$age <- paste0(round(mean(resonance[resonance$m16s_age_yr>=0.33 & resonance$m16s_age_yr<=0.67,]$m16s_age_yr)*12,0), " (",
                          round(min(resonance[resonance$m16s_age_yr>=0.33 & resonance$m16s_age_yr<=0.67,]$m16s_age_yr)*12,0), ", ",
                          round(max(resonance[resonance$m16s_age_yr>=0.33 & resonance$m16s_age_yr<=0.67,]$m16s_age_yr)*12,0), ")")

sum_stool_3 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_3, ms_id_3)
sum_stool_3$age <- paste0(round(mean(resonance[resonance$m16s_age_yr>=0.75 & resonance$m16s_age_yr<=1.58,]$m16s_age_yr)*12,0), " (",
                          round(min(resonance[resonance$m16s_age_yr>0.75 & resonance$m16s_age_yr<=1.58,]$m16s_age_yr)*12,0), ", ",
                          round(max(resonance[resonance$m16s_age_yr>=0.75 & resonance$m16s_age_yr<=1.58,]$m16s_age_yr)*12,0), ")")

sum_stool_4 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_4, ms_id_4)
sum_stool_4$age <- paste0(round(mean(resonance[resonance$m16s_age_yr>=1.67 & resonance$m16s_age_yr<=3.42,]$m16s_age_yr)*12,0), " (",
                          round(min(resonance[resonance$m16s_age_yr>=1.67 & resonance$m16s_age_yr<=3.42,]$m16s_age_yr)*12,0), ", ",
                          round(max(resonance[resonance$m16s_age_yr>=1.67 & resonance$m16s_age_yr<=3.42,]$m16s_age_yr)*12,0), ")")

sum_stool_resonance_2 <- rbind(sum_stool_1, sum_stool_2, sum_stool_3, sum_stool_4)
sum_stool_resonance_2$cohort <- "RESONANCE"

```

# Visualize fecal-fecal dissimilarity between child-mother and child-random mother
## Rarefied
```{r fig.width=12, fig.height=10}
sum_stool <- rbind(sum_stool_march, sum_stool_vdaart, sum_stool_resonance)
sum_stool <- sum_stool %>% group_by(cohort, beta, age, comparison) %>% mutate(size=n())
sum_stool$age <- paste0(sum_stool$age, "\nn=", sum_stool$size)

table(sum_stool$age)
sum_stool$age <- factor(sum_stool$age, levels=c("2 (1, 3)\nn=41", "3 (3, 6)\nn=23", "4 (2, 6)\nn=75", "5 (4, 8)\nn=25", 
                                                "12 (10, 15)\nn=97", "13 (10, 19)\nn=24", "32 (20, 41)\nn=13", "36 (34, 39)\nn=74", 
                                                "48 (47, 51)\nn=56"))


sum_stool %>%
  ggplot(aes(x=age, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age in Months, Mean (Range)", y="Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols=vars(cohort), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8,
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```
```{r fig.width=6, fig.height=10}
sum_stool %>%
  ggplot(aes(x=1, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs( y="Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols=vars(cohort),  scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8, 
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```
```{r fig.width=3, fig.height=10}
sum_stool %>%
  ggplot(aes(x=1, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs( y="Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8, 
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```

Save rarefied figure by cohort and child age, only show Jaccard and Bray-Curtis
```{r fig.width=12, fig.height=5}
f1 <-
  sum_stool[sum_stool$beta=="Jaccard" | sum_stool$beta=="Bray-Curtis",] %>%
  ggplot(aes(x=age, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age in Months, Mean (Range)", y="Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols=vars(cohort), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8,
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

f1
```

## Unrarefied
```{r fig.width=12, fig.height=4.5}
sum_stool_2 <- rbind(sum_stool_march_2, sum_stool_vdaart_2, sum_stool_resonance_2)
sum_stool_2 <- sum_stool_2 %>% group_by(cohort, beta, age, comparison) %>% mutate(size=n())
sum_stool_2$age <- paste0(sum_stool_2$age, "\nn=", sum_stool_2$size)

sum_stool_2$age <- factor(sum_stool_2$age, levels=c("2 (1, 3)\nn=41", "3 (3, 6)\nn=23", "4 (2, 6)\nn=75", "5 (4, 8)\nn=25", 
                                                "12 (10, 15)\nn=97", "13 (10, 19)\nn=24", "32 (20, 41)\nn=13", "36 (34, 39)\nn=74", 
                                                "48 (47, 51)\nn=56"))

sum_stool_2[sum_stool_2$beta=="Jaccard" | sum_stool_2$beta=="Bray-Curtis",] %>%
  ggplot(aes(x=age, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age in Months, Mean (Range)", y="Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols=vars(cohort), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8,
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```
```{r fig.width=6, fig.height=4.5}
sum_stool_2[sum_stool_2$beta=="Jaccard" | sum_stool_2$beta=="Bray-Curtis",] %>%
  ggplot(aes(x=1, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(y="Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols=vars(cohort),  scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8,
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```
```{r fig.width=3, fig.height=4.5}
sum_stool_2[sum_stool_2$beta=="Jaccard" | sum_stool_2$beta=="Bray-Curtis",] %>%
  ggplot(aes(x=1, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age (Months)", y="Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8,
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```



# WISC
```{r}
wisc <- read.csv(paste0(path,"Results/feast_11305_091123.csv"))

table(wisc$m16s_age_yr)
# possible cut off: 0.08-0.17, 0.5-0.92
```

## Extract infant ID, maternal vaginal ID, maternal vaginal ID
```{r}
i_id_1 <- wisc %>% filter(m16s_age_yr<=0.17) %>% select(sample.id)
mv_id_1 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==11305 & 
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_1$sample.id, 7, 15)))) %>% select(sample.id)

i_id_2 <- wisc %>% filter(m16s_age_yr>=0.5 & m16s_age_yr<=0.92) %>% select(sample.id)
mv_id_2 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==11305 & 
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_2$sample.id, 7, 15)))) %>% select(sample.id)
```

## Read in dissimilarity matrices
```{r message=FALSE, warning=FALSE}
# bray-curtis
bc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/11305/braycurtis.tsv") 
row_name <- bc$...1
bc <- bc[, c(-1)]
row.names(bc) <- row_name

# jaccard
jc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/11305/jaccard.tsv")
row_name <- jc$...1
jc <- jc[, c(-1)]
row.names(jc) <- row_name

# unweighted unifrac
uu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/11305/unweighted_unifrac.tsv")
row_name <- uu$...1
uu <- uu[, c(-1)]
row.names(uu) <- row_name

# weighted unifrac
wu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/11305/weighted_unifrac.tsv")
row_name <- wu$...1
wu <- wu[, c(-1)]
row.names(wu) <- row_name
```



## Create summary datasets for dissimilarity within pairs of mother-own child, mother-unrelated child
```{r message=FALSE}
## based on rarefied count table
sum_vaginal_1 <- combine_sum_diss(bc, jc, uu, wu, i_id_1, mv_id_1)
sum_vaginal_1$age <- paste0(round(mean(wisc[wisc$m16s_age_yr<=0.17,]$m16s_age_yr, na.rm = TRUE)*12,0), " (",
                          round(min(wisc[wisc$m16s_age_yr<=0.17,]$m16s_age_yr, na.rm = TRUE)*12,0), ", ",
                          round(max(wisc[wisc$m16s_age_yr<=0.17,]$m16s_age_yr, na.rm = TRUE)*12,0), ")")

sum_vaginal_2 <- combine_sum_diss(bc, jc, uu, wu, i_id_2, mv_id_2)
sum_vaginal_2$age <- paste0(round(mean(wisc[wisc$m16s_age_yr>=0.5 & wisc$m16s_age_yr<=0.92,]$m16s_age_yr, na.rm = TRUE)*12,0), " (",
                          round(min(wisc[wisc$m16s_age_yr>=0.5 & wisc$m16s_age_yr<=0.92,]$m16s_age_yr, na.rm = TRUE)*12,0), ", ",
                          round(max(wisc[wisc$m16s_age_yr>=0.5 & wisc$m16s_age_yr<=0.92,]$m16s_age_yr, na.rm = TRUE)*12,0), ")")

sum_vaginal_wisc <- rbind(sum_vaginal_1, sum_vaginal_2)
sum_vaginal_wisc$cohort <- "WISC"


## based on un-rarefied count table
asv_table <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/11305/11305_asv_table_classic.tsv")
  # this is newer asv table Justine created in May 2023

asv_table <- as.data.frame(t(asv_table)) 
  # rows are samples, columns are taxa
colnames(asv_table) <- asv_table[1, ]
  # assign colnames as ASV id
asv_table <- asv_table[-(1:2), ] # remove unnecessary rows
asv_table <- asv_table[-(nrow(asv_table)), ] # remove unnecessary rows

row_names <- row.names(asv_table)
asv_table <- as.data.frame(lapply(asv_table, as.integer))
asv_table <- as.matrix(asv_table)
row.names(asv_table) <- row_names

jc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "jaccard")))
bc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "bray")))


sum_vaginal_1 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_1, mv_id_1)
sum_vaginal_1$age <- paste0(round(mean(wisc[wisc$m16s_age_yr<=0.17,]$m16s_age_yr, na.rm = TRUE)*12,0), " (",
                          round(min(wisc[wisc$m16s_age_yr<=0.17,]$m16s_age_yr, na.rm = TRUE)*12,0), ", ",
                          round(max(wisc[wisc$m16s_age_yr<=0.17,]$m16s_age_yr, na.rm = TRUE)*12,0), ")")

sum_vaginal_2 <- combine_sum_diss(bc_2, jc_2, uu, wu, i_id_2, mv_id_2)
sum_vaginal_2$age <- paste0(round(mean(wisc[wisc$m16s_age_yr>=0.5 & wisc$m16s_age_yr<=0.92,]$m16s_age_yr, na.rm = TRUE)*12,0), " (",
                          round(min(wisc[wisc$m16s_age_yr>=0.5 & wisc$m16s_age_yr<=0.92,]$m16s_age_yr, na.rm = TRUE)*12,0), ", ",
                          round(max(wisc[wisc$m16s_age_yr>=0.5 & wisc$m16s_age_yr<=0.92,]$m16s_age_yr, na.rm = TRUE)*12,0), ")")

sum_vaginal_wisc_2 <- rbind(sum_vaginal_1, sum_vaginal_2)
sum_vaginal_wisc_2$cohort <- "WISC"

```

# Rochester
```{r}
rochester <- read.csv(paste0(path,"Results/feast_12601_091123.csv"))

table(rochester$m16s_age_yr)
# possible cut off: 0-0.17, 0.42-0.75, 0.83-1.5
```

## Extract infant ID, maternal vaginal ID, maternal vaginal ID
```{r}
i_id_1 <- rochester %>% filter(m16s_age_yr<=0.17) %>% select(sample.id)
mv_id_1 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==12601 & 
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_1$sample.id, 7, 15)))) %>% select(sample.id)

i_id_2 <- rochester %>% filter(m16s_age_yr>=0.42 & m16s_age_yr<=0.75) %>% select(sample.id)
mv_id_2 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==12601 & 
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_2$sample.id, 7, 15)))) %>% select(sample.id)

i_id_3 <- rochester %>% filter(m16s_age_yr>=0.83 & m16s_age_yr<=1.5) %>% select(sample.id)
mv_id_3 <- full_dta %>% filter(m16s_part_type=="biological mother" & CohortID==12601 & 
                                 (substr(ParticipantID, 1, 9) %in% as.list(substr(i_id_3$sample.id, 7, 15)))) %>% select(sample.id)
```


```{r}
min(full_dta[full_dta$sample.id %in% as.list(i_id_1$sample.id),]$m16s_slv138_97_depth)
length(unique(full_dta[full_dta$sample.id %in% as.list(i_id_1$sample.id),]$sample.id))

min(full_dta[full_dta$sample.id %in% as.list(i_id_2$sample.id),]$m16s_slv138_97_depth)
length(unique(full_dta[full_dta$sample.id %in% as.list(i_id_2$sample.id),]$sample.id))

min(full_dta[full_dta$sample.id %in% as.list(i_id_3$sample.id),]$m16s_slv138_97_depth)
length(unique(full_dta[full_dta$sample.id %in% as.list(i_id_3$sample.id),]$sample.id))

min(full_dta[full_dta$sample.id %in% as.list(mv_id_1$sample.id),]$m16s_slv138_97_depth)
length(unique(full_dta[full_dta$sample.id %in% as.list(mv_id_1$sample.id),]$sample.id))

min(full_dta[full_dta$sample.id %in% as.list(mv_id_2$sample.id),]$m16s_slv138_97_depth)
length(unique(full_dta[full_dta$sample.id %in% as.list(mv_id_2$sample.id),]$sample.id))

min(full_dta[full_dta$sample.id %in% as.list(mv_id_3$sample.id),]$m16s_slv138_97_depth)
length(unique(full_dta[full_dta$sample.id %in% as.list(mv_id_3$sample.id),]$sample.id))

min(full_dta[full_dta$CohortID==12601,]$m16s_slv138_97_depth)
```

## Read in dissimilarity matrices
Because many of IDs in matrices do not overlap with IDs contributed to analysis, these matrices are useless. 
I calculated Jaccard and Bray-Curtis based on ASV tables as below
```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# bray-curtis
bc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12601/braycurtis.tsv") 
row_name <- bc$...1
bc <- bc[, c(-1)]
row.names(bc) <- row_name

# jaccard
jc <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12601/jaccard.tsv")
row_name <- jc$...1
jc <- jc[, c(-1)]
row.names(jc) <- row_name

# unweighted unifrac
uu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12601/unweighted_unifrac.tsv")
row_name <- uu$...1
uu <- uu[, c(-1)]
row.names(uu) <- row_name

# weighted unifrac
wu <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12601/weighted_unifrac.tsv")
row_name <- wu$...1
wu <- wu[, c(-1)]
row.names(wu) <- row_name
```

## Create summary datasets for dissimilarity within pairs of mother-own child, mother-unrelated child
```{r message=FALSE}
## based on un-rarefied count table
asv_table <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12601/12601_asv_table_classic.tsv")
  # this is newer asv table Justine created in May 2023

asv_table <- as.data.frame(t(asv_table)) 
  # rows are samples, columns are taxa
colnames(asv_table) <- asv_table[1, ]
  # assign colnames as ASV id
asv_table <- asv_table[-(1:2), ] # remove unnecessary rows
asv_table <- asv_table[-(nrow(asv_table)), ] # remove unnecessary rows

row_names <- row.names(asv_table)
asv_table <- as.data.frame(lapply(asv_table, as.integer))
asv_table <- as.matrix(asv_table)
row.names(asv_table) <- row_names

jc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "jaccard")))
bc_2 <- as.data.frame(as.matrix(vegdist(asv_table, method = "bray")))


sum_vaginal_1 <- rbind(sum_diss(bc_2, i_id_1, mv_id_1), sum_diss(jc_2, i_id_1, mv_id_1))
sum_vaginal_1$age <- paste0(round(mean(rochester[rochester$m16s_age_yr<=0.17,]$m16s_age_yr)*12,0), " (",
                          round(min(rochester[rochester$m16s_age_yr<=0.17,]$m16s_age_yr)*12,0), ", ",
                          round(max(rochester[rochester$m16s_age_yr<=0.17,]$m16s_age_yr)*12,0), ")")

sum_vaginal_2 <- rbind(sum_diss(bc_2, i_id_2, mv_id_2), sum_diss(jc_2, i_id_2, mv_id_2))
sum_vaginal_2$age <- paste0(round(mean(rochester[rochester$m16s_age_yr>=0.42 & rochester$m16s_age_yr<=0.75,]$m16s_age_yr)*12,0), " (",
                          round(min(rochester[rochester$m16s_age_yr>=0.42 & rochester$m16s_age_yr<=0.75,]$m16s_age_yr)*12,0), ", ",
                          round(max(rochester[rochester$m16s_age_yr>=0.42 & rochester$m16s_age_yr<=0.75,]$m16s_age_yr)*12,0), ")")

sum_vaginal_3 <- rbind(sum_diss(bc_2, i_id_3, mv_id_3), sum_diss(jc_2, i_id_3, mv_id_3))
sum_vaginal_3$age <- paste0(round(mean(rochester[rochester$m16s_age_yr>=0.83 & rochester$m16s_age_yr<=1.5,]$m16s_age_yr)*12,0), " (",
                          round(min(rochester[rochester$m16s_age_yr>=0.83 & rochester$m16s_age_yr<=1.5,]$m16s_age_yr)*12,0), ", ",
                          round(max(rochester[rochester$m16s_age_yr>=0.83 & rochester$m16s_age_yr<=1.5,]$m16s_age_yr)*12,0), ")")

sum_vaginal_rochester_2 <- rbind(sum_vaginal_1, sum_vaginal_2, sum_vaginal_3)
sum_vaginal_rochester_2$cohort <- "Rochester"


## based on rarefied count table
asv_table_r <- rrarefy(asv_table, 1000) # rarefy at 1000

jc <- as.data.frame(as.matrix(vegdist(asv_table_r, method = "jaccard")))
bc <- as.data.frame(as.matrix(vegdist(asv_table_r, method = "bray")))

sum_vaginal_1 <- rbind(sum_diss(bc, i_id_1, mv_id_1), sum_diss(jc, i_id_1, mv_id_1))
sum_vaginal_1$age <- paste0(round(mean(rochester[rochester$m16s_age_yr<=0.17,]$m16s_age_yr)*12,0), " (",
                          round(min(rochester[rochester$m16s_age_yr<=0.17,]$m16s_age_yr)*12,0), ", ",
                          round(max(rochester[rochester$m16s_age_yr<=0.17,]$m16s_age_yr)*12,0), ")")

sum_vaginal_2 <- rbind(sum_diss(bc, i_id_2, mv_id_2), sum_diss(jc, i_id_2, mv_id_2))
sum_vaginal_2$age <- paste0(round(mean(rochester[rochester$m16s_age_yr>=0.42 & rochester$m16s_age_yr<=0.75,]$m16s_age_yr)*12,0), " (",
                          round(min(rochester[rochester$m16s_age_yr>=0.42 & rochester$m16s_age_yr<=0.75,]$m16s_age_yr)*12,0), ", ",
                          round(max(rochester[rochester$m16s_age_yr>=0.42 & rochester$m16s_age_yr<=0.75,]$m16s_age_yr)*12,0), ")")

sum_vaginal_3 <- rbind(sum_diss(bc, i_id_3, mv_id_3), sum_diss(jc, i_id_3, mv_id_3))
sum_vaginal_3$age <- paste0(round(mean(rochester[rochester$m16s_age_yr>=0.83 & rochester$m16s_age_yr<=1.5,]$m16s_age_yr)*12,0), " (",
                          round(min(rochester[rochester$m16s_age_yr>=0.83 & rochester$m16s_age_yr<=1.5,]$m16s_age_yr)*12,0), ", ",
                          round(max(rochester[rochester$m16s_age_yr>=0.83 & rochester$m16s_age_yr<=1.5,]$m16s_age_yr)*12,0), ")")

sum_vaginal_rochester <- rbind(sum_vaginal_1, sum_vaginal_2, sum_vaginal_3)
sum_vaginal_rochester$cohort <- "Rochester"

```


# Visualize vaginal-fecal dissimilarity between child-mother and child-random mother
## Rarefied
```{r fig.width=12, fig.height=10}
sum_vaginal <- rbind(sum_vaginal_march, sum_vaginal_wisc, sum_vaginal_rochester)
sum_vaginal <- sum_vaginal %>% group_by(cohort, beta, age, comparison) %>% mutate(size=n())
sum_vaginal$age <- paste0(sum_vaginal$age, "\nn=", sum_vaginal$size)

table(sum_vaginal$age)
sum_vaginal$age <- factor(sum_vaginal$age, levels=c("1 (0, 2)\nn=48", "2 (1, 2)\nn=127", "3 (3, 6)\nn=23", "6 (5, 9)\nn=87", "9 (6, 11)\nn=125",
                                                    "12 (10, 18)\nn=88"))

sum_vaginal %>%
  ggplot(aes(x=age, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age in Months, Mean (Range)", y = "Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols=vars(cohort), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8,
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 
```
```{r}
sum_vaginal[sum_vaginal$cohort=="Rochester",] %>%
  group_by(cohort, beta, age, comparison) %>%
  summarise(mean=mean(values))
```

```{r fig.width=6, fig.height=10}
sum_vaginal %>%
  ggplot(aes(x=1, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age (Months)", y = "Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols = vars(cohort), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8, 
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```
```{r fig.width=3, fig.height=5}
sum_vaginal %>%
  ggplot(aes(x=1, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age (Months)", y = "Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8, 
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```

Save rarefied figure by cohort and child age, only show Jaccard and Bray-Curtis
```{r fig.width=12, fig.height=5}
f2 <-
  sum_vaginal[sum_vaginal$beta=="Jaccard" | sum_vaginal$beta=="Bray-Curtis",] %>%
  ggplot(aes(x=age, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age in Months, Mean (Range)", y = "Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols=vars(cohort), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.6,
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

f2
```

## Unrarefied
```{r fig.width=12, fig.height=5}
sum_vaginal_2 <- rbind(sum_vaginal_march_2, sum_vaginal_wisc_2, sum_vaginal_rochester_2)
sum_vaginal_2 <- sum_vaginal_2 %>% group_by(cohort, beta, age, comparison) %>% mutate(size=n())
sum_vaginal_2$age <- paste0(sum_vaginal_2$age, "\nn=", sum_vaginal_2$size)

table(sum_vaginal_2$age)
sum_vaginal_2$age <- factor(sum_vaginal_2$age, levels=c("1 (0, 2)\nn=48", "2 (1, 2)\nn=127", "3 (3, 6)\nn=23", "6 (5, 9)\nn=87", "9 (6, 11)\nn=125",
                                                    "12 (10, 18)\nn=88"))

sum_vaginal_2[sum_vaginal_2$beta=="Jaccard" | sum_vaginal_2$beta=="Bray-Curtis",] %>%
  ggplot(aes(x=age, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age in Months, Mean (Range)", y = "Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols=vars(cohort), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8,
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```
```{r fig.width=6, fig.height=5}
sum_vaginal_2[sum_vaginal_2$beta=="Jaccard" | sum_vaginal_2$beta=="Bray-Curtis",] %>%
  ggplot(aes(x=1, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age (Months)", y = "Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), cols = vars(cohort), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8, 
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```
```{r fig.width=3, fig.height=5}
sum_vaginal_2[sum_vaginal_2$beta=="Jaccard" | sum_vaginal_2$beta=="Bray-Curtis",] %>%
  ggplot(aes(x=1, y = values, fill=comparison)) +
  geom_split_violin(trim = TRUE, alpha = .3) +  labs(x = "Child's Age (Months)", y = "Dissimilarity", fill="Pair") + 
  geom_boxplot(width = .15, alpha = .5, position = position_dodge(.2)) +
  facet_grid(rows = vars(beta), scales = "free", space = "free_x") +
  scale_fill_viridis_d(direction=-1) +
  theme_bw() + 
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14), strip.text = element_text(size=14), 
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.position = "bottom", legend.text = element_text(size=14), legend.title = element_text(size=14)) +
  stat_compare_means(method = "wilcox.test", label = "p.signif", size=7, color="red", vjust=0.8, 
                     symnum.args=list(cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", "+", ""))) 

```

# Make final figure
```{r fig.width=12, fig.height=10}
p <- 
  ggarrange(NULL, f1 + theme(strip.background = element_rect(fill="white")), NULL, f2+theme(strip.background = element_rect(fill="white")), 
            nrow=4, 
          common.legend = TRUE, legend = "bottom", heights=c(0.08, 1, 0.08, 1),
          labels = c("A. Maternal Fecal vs Child Fecal Microbiota", "", "B. Maternal Vaginal vs Child Fecal Microbiota", ""), 
          font.label = list(size = 16), hjust=0)
p

    ggsave(p, filename = "Similarity_103123.png",
       path =  paste0(path, "Figures"),
       height = 10, width = 12, units = "in", dpi = 500)

```