---
title: "Estimate the Overlap between Maternal Vaginal and Fecal microbiota in MARCH"
author: "Tiange Liu"
date: "`r Sys.Date()`"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/USERS/TiangeLiu/OUTPUT/Results",  
    )
  })

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(qiime2R)
library(FEAST)
library(data.table)
library(table1)
library(tidyr)
library(readr)
library(tidyverse)
source("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/USERS/TiangeLiu/ANALYSIS/Codes/OIF_Function.r")
```


# Read in participant data, explore, and modify
```{r message=FALSE}
march <- read.csv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/12902/MARCH_paired_54.csv")
  # these 54 dyads include additional 4 infants: set 6, 5-6 months when samples were collected

table(march$m16s_part_type)
march$participant_type <- as.factor(march$m16s_part_type)
march$participant_type <- factor(march$participant_type, levels=c("1", "2"), labels=c("biological mother", "child"))

# for this analysis, only keep mothers
march <- filter(march, participant_type=="biological mother")

table(march$m16s_spec_type)
march$body_site <- as.factor(march$m16s_spec_type)
march$body_site <- factor(march$m16s_spec_type, levels=c("1", "7"), labels=c("stool", "vaginal"))

march$paired_ID <- substr(march$ParticipantID, 1, 9)
```

# Read in covariates 
```{r message=FALSE}
master_dt <- read.csv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/OIF_master_SQL_072823_v3.csv")
master_dt[5500:5505, c("ParticipantID.x", "ParticipantID.y")]
```

   
# FEAST
## Prepare ASV table for FEAST
```{r message=FALSE}
asv_table <- read_tsv("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/DATA/DERIVED/20230518/01.CohortMerge/03.cohort_asv_data/12902/12902_asv_table_classic.tsv")
  # this is newer asv table Justine created in May 2023

#asv_table$type 
#asv_table$data[1:10, 1:10]
asv_table <- as.data.frame(t(asv_table)) 
  # rows are samples, columns are taxa
colnames(asv_table) <- asv_table[1, ]
  # assign colnames as ASV id
asv_table <- asv_table[-(1:2), ] # remove unnecessary rows

# only keep paired maternal samples
asv_table_paired <- asv_table %>% filter(row.names(asv_table) %in% as.list(march$sample.id))
row_names <- row.names(asv_table_paired)

str(asv_table_paired[1:5, 1:5]) # counts are not integer now
asv_table_paired <- as.data.frame(lapply(asv_table_paired, as.integer)) 
  # transform counts to integer, which is required by FEAST

row.names(asv_table_paired) <- row_names

asv_table_paired <- asv_table_paired[order(row.names(asv_table_paired)), ]
asv_table_paired <- as.matrix(asv_table_paired)
  # this will be the input ASV table for FEAST
```

## Prepare input meta table for FEAST
```{r message=FALSE, warning=FALSE}
meta_dt <- march[, c("sample.id", "paired_ID", "body_site", "participant_type")]
meta_dt <- meta_dt[order(meta_dt$sample.id), ] 

meta_dt$SourceSink <- ifelse(meta_dt$body_site=="vaginal", "Sink", "Source")
  # maternal vaginal microbiota = sink, maternal fecal microbiota = source

# per requirement of FEAST, each pair should have same id
meta_dt$id <- 1
n <- 1
for(i in 2:nrow(meta_dt)){
  if(meta_dt$SourceSink[i]=="Source"){
    meta_dt$id[i] <- n+1
    n <- n+1
  }else{
    meta_dt$id[i] <- meta_dt$id[i-1]
  }
}

meta_dt$Env <- paste(meta_dt$body_site, meta_dt$id)

row.names(meta_dt) <- meta_dt$sample.id
meta_dt <- meta_dt[, c("sample.id", "Env", "SourceSink", "id")]
meta_dt$id <- as.integer(meta_dt$id) # integer is required by FEAST
meta_dt <- meta_dt[order(meta_dt$sample.id), ]
meta_dt <- as.data.frame(meta_dt)
meta_dt <- meta_dt[, c("Env", "SourceSink", "id")]
  # this will be the input metadata for FEAST

# check if rownames match (required by FEAST)
common.sample.ids <- intersect(rownames(meta_dt), rownames(asv_table_paired))
length(common.sample.ids) # 108
```


## Run FEAST
```{r eval=FALSE, warning=FALSE, include=FALSE}
set.seed(91423)

FEAST_output <- FEAST(C = asv_table_paired, metadata = meta_dt, different_sources_flag = 1, 
                      dir_path="Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/USERS/TiangeLiu/OUTPUT/Results",   
                      outfile="feast_12902_091423_overlap")
  # default: 1000 EM iteration; 
  # rarefied at minimal sequencing depth within each group of sink and corresponding source
```


## Modify FEAST output for visulization
```{r message=FALSE}
output <- read.table("Z:/DAC_AnalysisProjects/10_Analysis_Proposals/EC0556A/USERS/TiangeLiu/OUTPUT/Results/feast_12902_091423_overlap_source_contributions_matrix.txt")

# select columns containing results when mother stool is the source
output_stool <- output %>% select(contains("stool"))


# generate a new output 
output_full <- as.data.frame(row.names(output))
output_full$stool <- rowSums(output_stool, na.rm = TRUE) * ifelse(rowSums(is.na(output_stool)) == ncol(output_stool), NA, 1) * 100
output_full$unknown <- output$Unknown * 100
  # multiply all estimated contributions by 100 to get the proportion (%)

colnames(output_full)[1] <- "sample.id"
output_full$sample.id <- substr(output_full$sample.id, 1,30)
output_full$ParticipantID <- substr(output_full$sample.id, 7, 17)


# add covariates
output_full <- merge(output_full, master_dt[, -c(1)], by.x="ParticipantID", by.y="ParticipantID.y", all=FALSE)

# add age
output_full <- merge(output_full, march[, c("sample.id", "m16s_age_yr", "m16s_coll_tmst")], by="sample.id")


# re-label some variables
output_full$birth_delivm <- as.factor(output_full$birth_delivm)
output_full$birth_delivm <- factor(output_full$birth_delivm, levels=c("1","2"), labels=c("Vaginal delivery", "Cesarean delivery"))
table(output_full$birth_delivm)

output_full$med_antib_pn <- as.factor(output_full$med_antib_pn)
output_full$med_antib_pn <- factor(output_full$med_antib_pn, levels=c("0", "1"), labels=c("No antibiotics", "Took antibiotics"))
table(output_full$med_antib_pn)

table(output_full$birth_delivm, output_full$med_antib_pn)

# exclude one mother with missing delivery mode
output_full <- filter(output_full, !is.na(output_full$birth_delivm)) # now 53 mothers left


# create a long dataset for plotting
output_full <- gather(output_full, type, prop, stool:unknown, factor_key=TRUE)
output_full$type <- factor(output_full$type, levels=c("unknown", "stool"),
                                labels=c("Other", "Fecal"))

```

## Visualize
```{r echo=FALSE, fig.height=6.5, fig.width=5.5}
# overall
p1 <- output_full %>% 
  group_by(type) %>%
  summarise(sum=mean(prop)) %>%
  ggplot(aes(fill=type, y=sum, x=1, label=round(sum, 1))) +
  geom_bar(stat="identity") + 
  geom_text(size = 5, position = position_stack(vjust = 0.5)) + 
  ylab("Mean Contribution of Source (%)") + xlab("Overall\n(n=53)") + labs(fill="Source") + ggtitle(" ") +
  theme_bw() +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), 
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        axis.text = element_text(size=13), axis.title = element_text(size=15, face="bold"),
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank())   +
  scale_fill_manual(values=color1_2) 

# by trimester of vaginal samples
p2 <- output_full %>%
  group_by(m16s_coll_tmst, type) %>%
  summarise(sum=mean(prop)) %>%
  ggplot(aes(fill=type, y=sum, x=factor(m16s_coll_tmst, levels=c("1", "2", "3"), labels=c("1\n(n=25)", "2\n(n=5)", "3\n(n=23)")), label=round(sum, 1))) + 
  geom_bar(stat="identity") +
  geom_text(size = 5, position = position_stack(vjust = 0.5)) + 
  ylab("Mean Contribution of Source (%)") + xlab(" ") + labs(fill="Source") + ggtitle("Trimester of Vaginal Samples") +
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.text.x = element_text(size=15, face="bold"), 
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        axis.title.x=element_blank(), axis.title = element_text(size=15, face="bold"),
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15),legend.direction = "vertical",
        panel.grid = element_blank()) +
  scale_fill_manual(values=color1_2)

p <-
  ggarrange(p1+theme(legend.text = element_text(size=15), legend.title = element_text(size=15), legend.direction = "horizontal"), 
          p2+theme(axis.ticks = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow=1, widths=c(0.2, 0.4), common.legend = TRUE, legend = "bottom")
p
```

### Only show fecal contribution
```{r echo=FALSE, fig.height=6.5, fig.width=5.5}
p1 <- 
  output_full[output_full$type=="Fecal", ] %>%
  ggplot(aes(y=prop, x=1)) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Contribution of Fecal Microbiota (%)")  + xlab("Overall\n(n=53)") + ggtitle(" ") +
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")



# by trimester of vaginal samples
p2 <- output_full[output_full$type=="Fecal", ] %>%
  ggplot(aes(y=prop, x=factor(m16s_coll_tmst, levels=c("1", "2", "3"), labels=c("1\n(n=25)", "2\n(n=5)", "3\n(n=23)")))) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Contribution of Fecal Microbiota (%)")  + xlab(" ") + ggtitle("Trimester of Vaginal Samples") +
  theme_bw() +
  theme(axis.text = element_text(size=13),  axis.ticks.x = element_blank(), axis.text.x = element_text(size=15, face="bold"),
        axis.title.x = element_blank(),
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")


p_feast <-
  ggarrange(p1, 
          p2+theme(axis.ticks = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow=1, widths=c(0.2, 0.4), common.legend = TRUE, legend = "bottom")
p_feast
```

# ASV overlapping
```{r echo=FALSE}
march[march$m16s_spec_type==7, ] %>%
  ggplot(aes(y=m16s_s138_obs_otus_1k, x=factor(m16s_coll_tmst, levels=c("1", "2", "3"), labels=c("1\n(n=25)", "2\n(n=5)", "3\n(n=23)")))) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Unique ASVs after Rarefaction")  + xlab(" ") + ggtitle("Trimester of Vaginal Samples") +
  theme_bw() +
  theme(axis.text = element_text(size=13),  axis.ticks.x = element_blank(), axis.text.x = element_text(size=15, face="bold"),
        axis.title.x = element_blank(),
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")


```


```{r echo=FALSE, message=FALSE}
asv_table_paired <- as.data.frame(asv_table_paired)

march <- filter(march, ParticipantID %in% as.list(output_full$ParticipantID))
  # exclude 1 mother with missing delivery mode

# Determine the sharing of ASVs between maternal vaginal and fecal samples 
# Split into 2 ASV tables
vagina_id <- filter(march, body_site=="vaginal")
asv_vagina <- asv_table_paired %>% filter(row.names(asv_table_paired) %in% as.list(vagina_id$sample.id))
asv_vagina <- asv_vagina[order(row.names(asv_vagina)), ]

stool_id <- filter(march, body_site=="stool")
asv_stool <- asv_table_paired %>% filter(row.names(asv_table_paired) %in% as.list(stool_id$sample.id))
asv_stool <- asv_stool[order(row.names(asv_stool)), ]

# Determine sharing (changed on 9/5/2023)
share_res <- asv_vagina

for(i in 1:ncol(share_res)){
  share_res[, i] <- ifelse((asv_vagina[,i]>0) & (asv_stool[,i]>0), 3, # shared between maternal vagina and stool
                           ifelse((asv_vagina[,i]>0) & (asv_stool[,i]==0), 1, # existed in maternal vagina only
                                  ifelse((asv_vagina[,i]==0) & (asv_stool[,i]>0), 2, # existed in maternal fecal only 
                                         4)))  # not existed
}

share_sum <- as.data.frame(vagina_id[, c("sample.id", "ParticipantID")])
share_sum <- share_sum[order(share_sum$ParticipantID), ]

share_sum <- share_sum %>%
  mutate(
    # count how many taxa are shared 
    both = rowSums(share_res==3), 
    vagina = rowSums(share_res==1),
    stool = rowSums(share_res==2)
  ) 


# Determine sharing after taking into account sequencing depth based on relative abundance
# presence = when relative abundance >= 0.05%
asv_vagina_rlt <- asv_vagina / rowSums(asv_vagina) * 100
asv_stool_rlt <- asv_stool / rowSums(asv_stool) * 100

share_res_rlt <- asv_vagina
for(i in 1:ncol(share_res_rlt)){
  share_res_rlt[, i] <- ifelse((asv_vagina_rlt[,i]>=0.05) & (asv_stool_rlt[,i]>=0.05), 3, # shared between maternal vagina and stool
                           ifelse((asv_vagina_rlt[,i]>=0.05) & (asv_stool_rlt[,i]<0.05), 1, # existed in maternal vagina only
                                  ifelse((asv_vagina_rlt[,i]<0.05) & (asv_stool_rlt[,i]>=0.05), 2, # existed in maternal fecal only 
                                         4)))  # not existed
}

share_sum_rlt <- as.data.frame(vagina_id[, c("sample.id", "ParticipantID")])
share_sum_rlt <- share_sum_rlt[order(share_sum_rlt$ParticipantID), ]

share_sum_rlt <- share_sum_rlt %>%
  mutate(
    # count how many taxa are shared 
    both = rowSums(share_res_rlt==3), 
    vagina = rowSums(share_res_rlt==1),
    stool = rowSums(share_res_rlt==2)
  ) 


# Determine sharing after taking into account sequencing depth 
# presence = when count >= 1count/1000sequences
sequence_vagina <- filter(march, body_site=="vaginal")
sequence_vagina <- select(sequence_vagina, c("sample.id", "m16s_slv138_97_depth"))
sequence_vagina <- sequence_vagina[order(sequence_vagina$'sample.id'),]

sequence_stool <- filter(march, body_site=="stool")
sequence_stool <- select(sequence_stool, c("sample.id", "m16s_slv138_97_depth"))
sequence_stool <- sequence_stool[order(sequence_stool$'sample.id'),]


share_res_sd <- asv_vagina
for(i in 1:ncol(share_res_sd)){
  share_res_sd[, i] <- ifelse((asv_vagina[,i]/sequence_vagina[,2]>=0.001) & (asv_stool[,i]/sequence_stool[,2]>=0.001), 3, 
                              # shared between maternal vagina and stool
                           ifelse((asv_vagina[,i]/sequence_vagina[,2]>=0.001) & (asv_stool[,i]/sequence_stool[,2]<0.001), 1, 
                                  # existed in maternal vagina only
                                  ifelse((asv_vagina[,i]/sequence_vagina[,2]<0.001) & (asv_stool[,i]/sequence_stool[,2]>=0.001), 2, 
                                         # existed in maternal fecal only 
                                         4)))  # not existed
}
share_sum_sd <- as.data.frame(vagina_id[, c("sample.id", "ParticipantID")])
share_sum_sd <- share_sum_sd[order(share_sum_sd$ParticipantID), ]

share_sum_sd <- share_sum_sd %>%
  mutate(
    # count how many taxa are shared 
    both = rowSums(share_res_sd==3), 
    vagina = rowSums(share_res_sd==1),
    stool = rowSums(share_res_sd==2)
  ) 



# Add covariates
share_sum <- merge(share_sum, master_dt[, -c(1)], by.x="ParticipantID", by.y="ParticipantID.y", all=FALSE)
share_sum_rlt <- merge(share_sum_rlt, master_dt[, -c(1)], by.x="ParticipantID", by.y="ParticipantID.y", all=FALSE )
share_sum_sd <- merge(share_sum_sd, master_dt[, -c(1)], by.x="ParticipantID", by.y="ParticipantID.y", all=FALSE )

share_sum$prop_v_f <- share_sum$both / (share_sum$vagina + share_sum$both) *100
share_sum_rlt$prop_v_f <- share_sum_rlt$both / (share_sum_rlt$vagina + share_sum_rlt$both) * 100
share_sum_sd$prop_v_f <- share_sum_sd$both / (share_sum_sd$vagina + share_sum_sd$both) * 100

# add age
junk <- filter(march, body_site=="vaginal")
share_sum <- merge(share_sum, junk[, c("ParticipantID", "m16s_coll_tmst")], by="ParticipantID")
share_sum_rlt <- merge(share_sum_rlt, junk[, c("ParticipantID", "m16s_coll_tmst")], by="ParticipantID")
share_sum_sd <- merge(share_sum_sd, junk[, c("ParticipantID", "m16s_coll_tmst")], by="ParticipantID")


# convert data from wide to long 
share_sum_2 <- gather(share_sum, type, n_count, both:stool, factor_key=TRUE)
share_sum_2$type <- factor(share_sum_2$type, 
                                  levels=c("vagina", "stool", "both"),
                                  labels=c("Vaginal only", "Fecal only", "Both"))

```

## Visualize: proportion of vaginal microbiota that are in common with fecal microbiota
### Present when count >=1
```{r echo=FALSE, fig.height=6.5, fig.width=5.5, message=FALSE, warning=FALSE}
p1 <- 
  share_sum %>%
  ggplot(aes(y=prop_v_f, x=1)) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Proportion of ASVs in Common with Fecal Microbiota (%)")  + xlab("Overall\n(n=53)") + ggtitle(" ") + ylim(0, 60) +
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")


# by trimester of vaginal samples
p2 <- 
  share_sum %>%
  ggplot(aes(y=prop_v_f, x=factor(m16s_coll_tmst, levels=c("1", "2", "3"), labels=c("1\n(n=25)", "2\n(n=5)", "3\n(n=23)")))) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Proportion of ASVs in Common with Fecal Microbiota (%)")  + xlab(" ") + ggtitle("Trimester of Vaginal Samples") + ylim(0, 60) +
  theme_bw() +
  theme(axis.text = element_text(size=13),  axis.ticks.x = element_blank(), axis.text.x = element_text(size=15, face="bold"),
        axis.title.x = element_blank(),
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")

p_prop <-
  ggarrange(p1, 
          p2+theme(axis.ticks = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow=1, widths=c(0.2, 0.4), common.legend = TRUE, legend = "bottom")
p_prop

share_sum %>%
  group_by(m16s_coll_tmst) %>%
  summarise(mean=mean(prop_v_f))
```

### Present when relative abundance >= 0.05%
```{r echo=FALSE, fig.height=6.5, fig.width=5.5, message=FALSE, warning=FALSE}
p1 <- 
  share_sum_rlt %>%
  ggplot(aes(y=prop_v_f, x=1)) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Proportion of ASVs in Common with Fecal Microbiota (%)")  + xlab("Overall\n(n=53)") + ggtitle(" ") + ylim(0, 60) +
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")


# by trimester of vaginal samples
p2 <- 
  share_sum_rlt %>%
  ggplot(aes(y=prop_v_f, x=factor(m16s_coll_tmst, levels=c("1", "2", "3"), labels=c("1\n(n=25)", "2\n(n=5)", "3\n(n=23)")))) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Proportion of ASVs in Common with Fecal Microbiota (%)")  + xlab(" ") + ggtitle("Trimester of Vaginal Samples") + ylim(0, 60) +
  theme_bw() +
  theme(axis.text = element_text(size=13),  axis.ticks.x = element_blank(), axis.text.x = element_text(size=15, face="bold"),
        axis.title.x = element_blank(),
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")

p_prop <-
  ggarrange(p1, 
          p2+theme(axis.ticks = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow=1, widths=c(0.2, 0.4), common.legend = TRUE, legend = "bottom")
p_prop

share_sum_rlt %>%
  group_by(m16s_coll_tmst) %>%
  summarise(mean=mean(prop_v_f))
```

### Present when count >= 1 count/ 1000 sequences
```{r echo=FALSE, fig.height=6.5, fig.width=5.5, message=FALSE, warning=FALSE}
p1 <- 
  share_sum_sd %>%
  ggplot(aes(y=prop_v_f, x=1)) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Proportion of ASVs in Common with Fecal Microbiota (%)")  + xlab("Overall\n(n=53)") + ggtitle(" ") + ylim(0, 60) +
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")


# by trimester of vaginal samples
p2 <- 
  share_sum_sd %>%
  ggplot(aes(y=prop_v_f, x=factor(m16s_coll_tmst, levels=c("1", "2", "3"), labels=c("1\n(n=25)", "2\n(n=5)", "3\n(n=23)")))) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Proportion of ASVs in Common with Fecal Microbiota (%)")  + xlab(" ") + ggtitle("Trimester of Vaginal Samples") + ylim(0, 60) +
  theme_bw() +
  theme(axis.text = element_text(size=13),  axis.ticks.x = element_blank(), axis.text.x = element_text(size=15, face="bold"),
        axis.title.x = element_blank(),
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")

p_prop <-
  ggarrange(p1, 
          p2+theme(axis.ticks = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow=1, widths=c(0.2, 0.4), common.legend = TRUE, legend = "bottom")
p_prop

share_sum_sd %>%
  group_by(m16s_coll_tmst) %>%
  summarise(mean=mean(prop_v_f))
```


## Visualize: absolute number of ASVs of vaginal microbiota that are in common with fecal microbiota
### Present when count >=1
```{r echo=FALSE, fig.height=6.5, fig.width=5.5, message=FALSE, warning=FALSE}
p1 <- 
  share_sum %>%
  ggplot(aes(y=both, x=1)) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Number of ASVs in Common with Fecal Microbiota")  + xlab("Overall\n(n=53)") + ggtitle(" ") +
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")

# by trimester of vaginal samples
p2 <- 
  share_sum %>%
  ggplot(aes(y=both, x=factor(m16s_coll_tmst, levels=c("1", "2", "3"), labels=c("1\n(n=25)", "2\n(n=5)", "3\n(n=23)")))) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Number of ASVs in Common with Fecal Microbiota")  + xlab(" ") + ggtitle("Trimester of Vaginal Samples") +
  theme_bw() +
  theme(axis.text = element_text(size=13),  axis.ticks.x = element_blank(), axis.text.x = element_text(size=15, face="bold"),
        axis.title.x = element_blank(),
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")

p_n <-
  ggarrange(p1, 
          p2+theme(axis.ticks = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow=1, widths=c(0.2, 0.4), common.legend = TRUE, legend = "bottom")
p_n

share_sum %>%
  group_by(m16s_coll_tmst) %>%
  summarise(mean=mean(both))
```

### Present when relative abundance >= 0.05%
```{r echo=FALSE, fig.height=6.5, fig.width=5.5, message=FALSE, warning=FALSE}
p1 <- 
  share_sum_rlt %>%
  ggplot(aes(y=both, x=1)) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Number of ASVs in Common with Fecal Microbiota")  + xlab("Overall\n(n=53)") + ggtitle(" ") +
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")

# by trimester of vaginal samples
p2 <- 
  share_sum_rlt %>%
  ggplot(aes(y=both, x=factor(m16s_coll_tmst, levels=c("1", "2", "3"), labels=c("1\n(n=25)", "2\n(n=5)", "3\n(n=23)")))) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Number of ASVs in Common with Fecal Microbiota")  + xlab(" ") + ggtitle("Trimester of Vaginal Samples") +
  theme_bw() +
  theme(axis.text = element_text(size=13),  axis.ticks.x = element_blank(), axis.text.x = element_text(size=15, face="bold"),
        axis.title.x = element_blank(),
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")

p_n <-
  ggarrange(p1, 
          p2+theme(axis.ticks = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow=1, widths=c(0.2, 0.4), common.legend = TRUE, legend = "bottom")
p_n

share_sum_rlt %>%
  group_by(m16s_coll_tmst) %>%
  summarise(mean=mean(both))
```

### Present when count >= 1 count/ 1000 sequences
```{r echo=FALSE, fig.height=6.5, fig.width=5.5, message=FALSE, warning=FALSE}
p1 <- 
  share_sum_sd %>%
  ggplot(aes(y=both, x=1)) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Number of ASVs in Common with Fecal Microbiota")  + xlab("Overall\n(n=53)") + ggtitle(" ") +
  theme_bw() +
  theme(axis.text = element_text(size=13), axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")

# by trimester of vaginal samples
p2 <- 
  share_sum_sd %>%
  ggplot(aes(y=both, x=factor(m16s_coll_tmst, levels=c("1", "2", "3"), labels=c("1\n(n=25)", "2\n(n=5)", "3\n(n=23)")))) +
  geom_boxplot(lwd=0.8, width=0.8, outlier.shape=NA) + geom_point(size=3, position=position_jitter(w = 0.1, h = 0.1), alpha=0.5) +
  ylab("Number of ASVs in Common with Fecal Microbiota")  + xlab(" ") + ggtitle("Trimester of Vaginal Samples") +
  theme_bw() +
  theme(axis.text = element_text(size=13),  axis.ticks.x = element_blank(), axis.text.x = element_text(size=15, face="bold"),
        axis.title.x = element_blank(),
        axis.title = element_text(size=15, face="bold"),
        plot.title = element_text(size=15, face="bold", hjust = 0.5), 
        legend.title = element_text(size=15, face="bold"), legend.text = element_text(size=15), legend.direction = "vertical",
        panel.grid = element_blank()) +
  stat_summary(fun.y="mean", shape=16, size=1, color="red")

p_n <-
  ggarrange(p1, 
          p2+theme(axis.ticks = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow=1, widths=c(0.2, 0.4), common.legend = TRUE, legend = "bottom")
p_n

share_sum_sd %>%
  group_by(m16s_coll_tmst) %>%
  summarise(mean=mean(both))
```




## Visualize by venn diagram
```{r echo=FALSE, fig.width=6, fig.height=6}
fecal_unique_avg = mean(share_sum$stool)
fecal_unique_avg
vaginal_unique_avg = mean(share_sum$vagina)
vaginal_unique_avg
common_avg = mean(share_sum$both)
common_avg

total_samples <- fecal_unique_avg + vaginal_unique_avg + common_avg
fecal_set <- 1:fecal_unique_avg
vaginal_set <- (fecal_unique_avg + 1):(fecal_unique_avg + vaginal_unique_avg)
common_set <- (fecal_unique_avg + vaginal_unique_avg + 1):total_samples

set_list <- list(
  Fecal = c(fecal_set, common_set),
  Vaginal = c(vaginal_set, common_set)
)

library(ggVennDiagram)

ggVennDiagram(set_list) 

```


